---
title: "Survey Snake Plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Survey Snake Plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 8
)
```

Snake plots display survey items as horizontal bands in a serpentine
layout — each row reverses direction and connects to the next through a
U-turn arc. This compact layout lets you compare many items at once while
preserving adjacency information such as inter-item correlations.

The package ships with three datasets from an experience sampling study of
university students (Neubauer & Schmiedek, 2024):

| Dataset | Rows | Description |
|---------|------|-------------|
| `ema_emotions` | 280 | Person-level means for 10 emotions (1--7) |
| `student_survey` | 280 | 34 items across 4 constructs, prefixed for faceting |
| `ema_beeps` | 500 | Beep-level timestamps + anger/happiness ratings |

```{r load}
library(snakeplot)

cols7 <- c("#C0392B", "#E67E22", "#F1C40F", "#95A5A6",
           "#2ECC71", "#3498DB", "#8E44AD")
labs7 <- c("1" = "Not at all", "2" = "Slightly", "3" = "Somewhat",
           "4" = "Moderate",   "5" = "Quite",    "6" = "Very",
           "7" = "Extremely")
```

## Daily EMA — ticks by time-of-day

When you have timestamped beep-level data, `var`, `day`, and `timestamp`
auto-pivot into one band per day. Ticks are positioned by time-of-day
with hour gridlines:

```{r angry_time, fig.height = 12}
survey_snake(ema_beeps, var = "angry", day = "day",
             timestamp = "start_time",
             colors = cols7, arc_fill = "none",
             level_labels = labs7,
             label_cex = 1.0, legend_cex = 0.85,
             title = "Anger — 14 days, ticks by time-of-day")
```

```{r happy_time, fig.height = 12}
survey_snake(ema_beeps, var = "happy", day = "day",
             timestamp = "start_time",
             colors = cols7, arc_fill = "none",
             level_labels = labs7,
             label_cex = 1.0, legend_cex = 0.85,
             title = "Happiness — 14 days, ticks by time-of-day")
```

## Daily EMA — distribution bars

The same daily data with `tick_shape = "bar"` shows how response
distributions shift across days:

```{r angry_bar, fig.height = 12}
survey_snake(ema_beeps, var = "angry", day = "day",
             tick_shape = "bar",
             colors = cols7, arc_fill = "none",
             level_labels = labs7,
             label_cex = 1.0, legend_cex = 0.85,
             title = "Anger — 14 days, distribution bars")
```

```{r happy_bar, fig.height = 12}
survey_snake(ema_beeps, var = "happy", day = "day",
             tick_shape = "bar",
             colors = cols7, arc_fill = "none",
             level_labels = labs7,
             label_cex = 1.0, legend_cex = 0.85,
             title = "Happiness — 14 days, distribution bars")
```

## Cross-sectional survey — distribution bars

Pass a data.frame directly — each column becomes a band. Sorting by
mean reorders items from lowest to highest average:

```{r bars, fig.height = 10}
survey_snake(ema_emotions, tick_shape = "bar", sort_by = "mean",
             colors = cols7, level_labels = labs7,
             label_cex = 1.0, legend_cex = 0.85,
             title = "Emotions — distribution bars, sorted by mean")
```

## Cross-sectional survey — tick lines

Individual responses shown as colored tick marks within proportional
zones:

```{r basic, fig.height = 10}
survey_snake(ema_emotions, sort_by = "mean",
             colors = cols7, level_labels = labs7,
             label_cex = 1.0, legend_cex = 0.85,
             title = "Emotions — tick lines, sorted by mean")
```

## Arc fill modes

The `arc_fill` parameter controls how U-turn arcs are colored:

- `"none"` (default) — two-tone split matching adjacent band shades
- `"correlation"` — tinted by inter-item Pearson *r*
- `"mean_prev"` — solid fill using the preceding band's shade
- `"blend"` — 50/50 average of adjacent band shades

```{r arc_corr, fig.height = 10}
survey_snake(ema_emotions, tick_shape = "line",
             arc_fill = "correlation", sort_by = "mean",
             colors = cols7, level_labels = labs7,
             label_cex = 1.0, legend_cex = 0.85,
             title = "Emotions — correlations at U-turns")
```

## Faceted plots

When column names share a prefix (e.g., `Emo_`, `Mot_`), `facet = TRUE`
auto-groups them into a multi-panel layout. The prefix is stripped from
labels automatically:

```{r facet, fig.width = 12, fig.height = 16}
survey_snake(student_survey,
             facet = TRUE, facet_ncol = 2L,
             tick_shape = "bar", sort_by = "mean",
             colors = cols7, level_labels = labs7,
             label_cex = 1.0, legend_cex = 0.85)
```

You can also provide explicit groups as a named list:

```{r facet_explicit, fig.width = 12, fig.height = 10}
survey_snake(student_survey,
             facet = list(
               Emotions   = grep("^Emo_", names(student_survey), value = TRUE),
               Motivation = grep("^Mot_", names(student_survey), value = TRUE)
             ),
             facet_ncol = 2L,
             tick_shape = "bar", sort_by = "mean",
             colors = cols7, level_labels = labs7,
             label_cex = 1.0, legend_cex = 0.85)
```

## Mean and median markers

A diamond shows the item mean; a dashed line shows the median:

```{r markers, fig.height = 10}
survey_snake(ema_emotions,
             tick_shape = "bar", sort_by = "mean",
             show_mean = TRUE, show_median = TRUE,
             colors = cols7, level_labels = labs7,
             label_cex = 1.0, legend_cex = 0.85,
             title = "Emotions — mean (diamond) and median (dashed)")
```
