---
title: "Survey Snake Plots"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Survey Snake Plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 8
)
```

Snake plots display survey items as horizontal bands in a serpentine
layout — each row reverses direction and connects to the next through a
U-turn arc. This compact layout lets you compare many items at once while
preserving adjacency information such as inter-item correlations.

The package ships with three datasets from an experience sampling study of
university students (Neubauer & Schmiedek, 2024), and 10 built-in color
palettes via `snake_palettes`:

```{r load}
library(snakeplot)

labs7 <- c("1" = "Not at all", "2" = "Slightly", "3" = "Somewhat",
           "4" = "Moderate",   "5" = "Quite",    "6" = "Very",
           "7" = "Extremely")
```

## Daily EMA — value distribution ticks

When you have beep-level data, `var` and `day` auto-pivot into one band
per day. Ticks are positioned within proportional zones by response level:

```{r daily_value, fig.height = 12}
survey_snake(ema_beeps, var = "angry", day = "day",
             colors = snake_palettes$ocean, level_labels = labs7,
             title = "Anger — 14 days, value distribution")
```

## Tick lines with correlation arcs

Individual responses as colored tick marks, with inter-item Pearson *r*
displayed at each U-turn arc:

```{r corr, fig.height = 10}
survey_snake(ema_emotions, tick_shape = "line",
             arc_fill = "correlation", sort_by = "mean",
             colors = snake_palettes$ocean, level_labels = labs7,
             title = "Emotions — correlations at U-turns")
```

## Dot plot with dark bands

Use `band_palette` for darker band shading. Dots with jitter show
individual responses:

```{r dots_dark, fig.height = 10}
survey_snake(ema_emotions, tick_shape = "dot", sort_by = "mean",
             colors = snake_palettes$ocean, level_labels = labs7,
             band_palette = c("#1a1228", "#1a2a42"),
             title = "Emotions — dots on dark bands")
```

## Faceted multi-construct

When column names share a prefix (e.g., `Emo_`, `Mot_`), `facet = TRUE`
auto-groups them into panels. Prefixes are stripped from labels:

```{r facet, fig.width = 12, fig.height = 16}
survey_snake(student_survey, facet = TRUE, facet_ncol = 2L,
             tick_shape = "bar", sort_by = "mean",
             colors = snake_palettes$ocean, level_labels = labs7)
```

## Daily EMA — distribution bars

The same daily data with `tick_shape = "bar"` shows how response
distributions shift across days. Use `bar_reverse = TRUE` to start
from the highest level:

```{r daily_bars, fig.height = 12}
survey_snake(ema_beeps, var = "happy", day = "day",
             tick_shape = "bar", bar_reverse = TRUE,
             colors = snake_palettes$ocean, level_labels = labs7,
             title = "Happiness — 14 days, distribution bars")
```

## Activity timelines

`activity_snake()` shows daily event timelines — rug ticks or duration
blocks on a dark ribbon:

```{r activity, fig.height = 8}
set.seed(42)
days <- c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")
d <- data.frame(
  day      = rep(days, each = 40),
  start    = round(runif(280, 360, 1400)),
  duration = 0
)
activity_snake(d)
```

## Sequence snake

`sequence_snake()` displays a state sequence as colored blocks flowing
through the serpentine layout — each block is one time point colored by
its state:

```{r sequence_snake, fig.height = 7}
set.seed(42)
verbs <- c("Read", "Write", "Discuss", "Listen",
           "Search", "Plan", "Code", "Review")
seq75 <- character(0)
while (length(seq75) < 75) {
  seq75 <- c(seq75, rep(sample(verbs, 1), sample(1:4, 1)))
}
seq75 <- seq75[seq_len(75)]
sequence_snake(seq75, title = "75-step learning sequence")
```

## Timeline snake

`timeline_snake()` takes a 3-column data.frame (role, start, end) and
auto-generates monthly blocks, transition labels, and band year labels:

```{r timeline_snake}
career <- data.frame(
  role  = c("Intern", "Junior Dev", "Mid Dev",
            "Senior Dev", "Tech Lead", "Architect"),
  start = c("2015-01", "2015-07", "2017-01",
            "2019-07", "2022-07", "2024-01"),
  end   = c("2015-06", "2016-12", "2019-06",
            "2022-06", "2023-12", "2024-12")
)
timeline_snake(career,
               title = "Software Engineer — Career Path (2015-2024)")
```

## Survey sequence

`survey_sequence()` renders 100% stacked horizontal bars in a
serpentine layout:

```{r survey_seq}
survey_sequence(ema_emotions, colors = snake_palettes$ocean)
```

## Sequential distribution

`sequential_dist()` is a monochrome variant of `survey_sequence()`:

```{r seq_dist}
sequential_dist(ema_emotions)
```

## Line snake

`line_snake()` draws a continuous intensity line winding through bands:

```{r line_snake, fig.height = 6}
set.seed(42)
hours <- seq(0, 1440, by = 10)
d_line <- data.frame(
  day   = rep(c("Mon", "Tue", "Wed", "Thu", "Fri"), each = length(hours)),
  time  = rep(hours, 5),
  value = sin(rep(hours, 5) / 1440 * 4 * pi) * 50 + 50 +
          rnorm(5 * length(hours), 0, 8)
)
line_snake(d_line, fill_color = "#e74c3c")
```

## Built-in palettes

10 palettes are available via `snake_palettes` or `snake_palette()`:

```{r palette_names}
names(snake_palettes)
```

Use any palette by name:

```r
survey_snake(ema_emotions, colors = snake_palettes$earth, tick_shape = "bar")
snake_palette("sunset", n = 5)  # interpolate to 5 colors
```

## Mean and median markers

A diamond shows the item mean; a dashed line shows the median:

```{r markers, fig.height = 10}
survey_snake(ema_emotions, tick_shape = "bar", sort_by = "mean",
             show_mean = TRUE, show_median = TRUE,
             colors = snake_palettes$ocean, level_labels = labs7,
             title = "Emotions — mean (diamond) and median (dashed)")
```
